---
title: "Assignment 3"
author: "Tavish Boyle"
date: "September 24, 2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
titanic = read.csv("titanic.csv")
```

## Assignment 3



Steps preparing the data set "titanic_omit". 

```{r titanic, echo=TRUE}
titanic$Survived = factor(titanic$Survived)
```

#1.	Check if there is any missing data
```{r, echo = TRUE}
sum(is.na(titanic$Age))

```

#2.Replace missing values by the variable's average  for numeric variables
```{r, echo = TRUE}
titanic$Age[is.na(titanic$Age)] = mean(titanic$Age, na.rm = TRUE)
```

#3.	Remove the rows that have missing categorical values 
```{r, echo = TRUE}
titanic_omit <- na.omit(titanic)
```

#4.	Remove column names, ID and unnecessary columns 
```{r, echo = TRUE}
titanic_omit$Name = NULL
titanic_omit$PassengerId = NULL
titanic_omit$Ticket = NULL
titanic_omit$SibSp = NULL
titanic_omit$Cabin = NULL
```

#5.	Calculate the mean age
```{r, echo = TRUE}
mean(titanic_omit$Age)
```

#6.	Calculate the mean age of female passengers
```{r, echo = TRUE}
mean(titanic_omit[titanic_omit$Sex == "female",]$Age)
```

#7.	Calculate the median fare of the passengers in Class 1
```{r, echo = TRUE}
median(titanic_omit[titanic_omit$Pclass == 1,]$Fare)
```

#8.	Calculate the median fare of the female passengers that are not in Class 1
```{r, echo = TRUE}
median(titanic_omit[titanic_omit$Pclass != 1,]$Fare)
```

#9.	Calculate the median age of survived passengers who are female and Class 1 or Class 2
```{r, echo = TRUE}
median(titanic_omit[titanic_omit$Pclass == 1 | titanic_omit$Pclass == 2 & titanic_omit$Survived == 1 & titanic_omit$Sex == "female",]$Age)
```

#10. Calculate the mean fare of female teenagers survived passengers
```{r, echo = TRUE}
mean(titanic_omit[titanic_omit$Age < 20 & titanic_omit$Survived == 1 & titanic_omit$Sex == "female",]$Fare)
```

#11.	Calculate the mean fare of female teenagers survived passengers for each class
```{r, echo = TRUE}
mean(titanic_omit[titanic_omit$Age < 20 & titanic_omit$Survived == 1 & titanic_omit$Sex == "female" & titanic_omit$Pclass == 1,]$Fare)
```

```{r, echo = TRUE}
mean(titanic_omit[titanic_omit$Age < 20 & titanic_omit$Survived == 1 & titanic_omit$Sex == "female" & titanic_omit$Pclass == 2,]$Fare)
```

```{r, echo = TRUE}
mean(titanic_omit[titanic_omit$Age < 20 & titanic_omit$Survived == 1 & titanic_omit$Sex == "female" & titanic_omit$Pclass == 3,]$Fare)
``` 

#12.	Calculate the ratio of Survived and not Survived for passengers who are who pays more than the average fare 
```{r, echo = TRUE}
(nrow(titanic_omit[titanic_omit$Fare >= median(titanic_omit$Fare) & titanic_omit$Survived == 1,])) / (nrow(titanic_omit[titanic_omit$Fare >= median(titanic_omit$Fare) & titanic_omit$Survived == 0,]))
```

#13.	Add column that standardizes the fare (subtract the mean and divide by standard deviation) and name it sfare
```{r, echo = TRUE}
stdevfare = sd(titanic_omit$Fare)
avgfare = mean(titanic_omit$Fare)
titanic_omit$Sfare <- (titanic_omit$Fare - avgfare) / stdevfare
```

#14.	Plot the bar chart of two variables: Sex and Survived for passengers younger than the average age 
```{r, echo = TRUE}
ggplot(titanic_omit)+geom_bar(mapping = aes(x = titanic_omit$Survived == 1 & titanic_omit$Age <= mean(titanic_omit$Age), fill = Sex))
```

#15.	Plot the density of age of passengers who survived and pay more than the average fare
```{r, echo = TRUE}
ggplot(titanic_omit)+geom_density(mapping = aes(x = titanic_omit$Survived == 1 & titanic_omit$Fare >= mean(titanic_omit$Fare)))
```

#16.	Plot two density curves in one graph of passengers embarked in Southampton and Queenstown
```{r, echo = TRUE}
ggplot(titanic_omit)+geom_density(mapping = aes(x = titanic_omit$Embarked == "S", color = "Southampton")) + geom_density(mapping = aes(x = titanic_omit$Embarked == "Q", color = "Queenstown"))
```

#17.	Add categorical variable named "cfare" that takes value "cheap" for passengers paying less the average fare and takes value "expensive" for passengers paying more than the average fare.  
```{r, echo = TRUE}
titanic_omit$cfare <- ifelse(titanic_omit$Fare<34.69, "Cheap", "Expensive")
```

#18.	Add categorical variable named "cage" that takes value 0 for age 0-10, 1 for age 10-20, 2 for age 20-30, and so on 
```{r, echo = TRUE}

titanic_omit$cage[titanic_omit$Age<=10] <- 0
titanic_omit$cage[titanic_omit$Age>10 & titanic_omit$Age<=20] <- 1
titanic_omit$cage[titanic_omit$Age>20 & titanic_omit$Age<=30] <- 2
titanic_omit$cage[titanic_omit$Age>30 & titanic_omit$Age<=40] <- 3
titanic_omit$cage[titanic_omit$Age>40 & titanic_omit$Age<=50] <- 4
titanic_omit$cage[titanic_omit$Age>50 & titanic_omit$Age<=60] <- 5
titanic_omit$cage[titanic_omit$Age>60 & titanic_omit$Age<=70] <- 6
titanic_omit$cage[titanic_omit$Age>70 & titanic_omit$Age<=80] <- 7 
```

#19.	Plot bar chart of two variables cage and Survived for passengers paying more than the average 
```{r, echo = TRUE}
ggplot(titanic_omit[titanic_omit$Fare < mean(titanic_omit$Fare),]) + geom_bar(mapping = aes(x = cage, fill = Survived))
```

#20.	Split the bar chart in 18 by Pclass.
```{r, echo = TRUE}
ggplot(titanic_omit[titanic_omit$Fare < mean(titanic_omit$Fare),]) + geom_bar(mapping = aes(x = cage, fill = Survived)) + facet_wrap(~Pclass)
```

