---
title: "Assignment 9"
author: "Tavish Boyle"
date: "October 13, 2018"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
titanic = read.csv("train.csv")
# Remove some columns
titanic$PassengerId =  NULL
titanic$Ticket =  NULL
titanic$Name = NULL
titanic$Cabin = NULL

# Correct variables' types
titanic$Survived = factor(titanic$Survived)
titanic$Pclass = factor(titanic$Pclass)

# Handle missing values

levels(titanic$Embarked) = c("S","C","Q","S")

```

Function 1
Write a function that inputs a dataset (data frame) and outputs the number of missing values in each variables  
```{r, include=TRUE}
F1 <- function(df){
  sapply(df, function(df) sum(is.na(df)))
}

F1(titanic)
```

Function 2
Write a function that inputs a dataset and outputs the names of variables with missing values.
```{r, include=TRUE}
F2 <- function(df){
  sapply(df, function(df) sum(is.na(df)))
}

F2(titanic)
```

Function 3
Write a function that inputs a dataset and outputs (1) the names of variables with missing values and (2) the corresponding number of missing values in each variables.

```{r, include=TRUE}
F3 <- function(df){
  sapply(df, function(df) sum(is.na(df)))
}
F3(titanic)
```

Function 4
Write a function that inputs a dataset and output a dataset with missing values handled as follows:
-	Missing values of numeric variables are replaced by the means of the non-missing values in the variables 
-	Missing values of categorical variables are replaced by the most frequent category in the variables

```{r, include=TRUE}
F4 <- function(df){
  for (i in 1:ncol(df)){
    if (is.numeric(df[,i])){
      df[,i][is.na(df[,i])]=mean(df[,i], na.rm=TRUE)
    }else{
      levels=unique(df[,i])
      df[,i][is.na(df[,i])]=levels[which.max(tabulate(match(df[,i], levels)))]
    }
  }
  return (head(df))
}
F4(titanic)
```

Function 5
Write a function that input a dataset and output (1) sub-dataset containing categorical variables and (2) sub-dataset containing numeric variables.
```{r, include = TRUE}
F5 <-  function(df){
  num <- sapply(df, is.numeric)
  numeric <- data.frame(df[,num])
  print(numeric)
  
  cat <- sapply(df, is.factor)
  categorical <- data.frame(df[,cat])
  print(categorical)
}
F5(titanic)
```

Function 6
Write a function that inputs a dataset and output density curves of all numeric variables.
```{r, include = TRUE}
F6 <- function(x){
  for (i in 1:ncol(x)){
    if (is.numeric(x[,i])){
      print(ggplot(data=x)+geom_density(mapping=aes(x=x[,i]))+xlab(names(x)[i]))
    }
  }
}
F6(titanic)
```

Function 7
Write a function that inputs a dataset and output bar charts of all categorical variables.
```{r, include=TRUE}
F7<- function(x){
  for (i in 1:ncol(x)){
    if (is.factor(x[,i])){
      print(ggplot(data=x)+geom_bar(mapping=aes(x=x[,i]))+xlab(names(x)[i]))
    }
  }
}
F7(titanic)
```

Function 8
Write a function that inputs a dataset and two number i, j and outputs bar charts of column i and column j.  If either i or j are not categorical variables, output a message saying so.
```{r, inlcude=TRUE}
F8 <- function(x, i, j){
 if(is.numeric(x[,i]) | is.numeric(x[,j])){
    print("At least one of the columns is not categorical")
  }else{
    print(ggplot(data=x)+geom_bar(mapping=aes(x=x[,i],fill=x[,j]),position="dodge")+labs(x=names(x)[i],fill=names(x)[j]))
  }
}
F8(titanic, 3, 8)
```

Function 9
Write a function that inputs a dataset and two number i, j and outputs the scatter plot of column i and column j. If either i or j are not numeric variables, output a message saying so.
```{r, include=TRUE}
F9 <- function(x, i, j){
  if(!is.numeric(x[,i]) | !is.numeric(x[,j])){
    print("At least one of the columns is not numeric")
  }else{
    print(ggplot(data=x)+geom_point(mapping=aes(x=x[,i],y=x[,j]))+labs(x=names(x)[i],y=names(x)[j]))
  }
}
F9(titanic, 7, 4)
```

Function 10
Write a function that inputs a dataset and two number i, j and output the density curves of column i partitioning on categories of column j. If the two columns are not suitable for the graph, print out the message saying so. 
```{r, include=TRUE}
F10 <- function(x, i, j){
  if(!is.numeric(x[,i])){
    print("The input variable is not numeric")
  }else if(is.numeric(x[,j])){
    print("The partitioning variable is not categorical")
  }else{
    print(ggplot(data=x)+geom_density(mapping=aes(x=x[,i],color=x[,j]))+labs(x=names(x)[i],color=names(x)[j]))
  }
}
F10(titanic,7,3)
```


Function 11
Write a function that inputs a dataset and output all possible bar charts of pair of variables in the dataset. 
```{r, include=TRUE}
F11 <- function(x){
  for (i in 1:ncol(x)){
      if (!is.numeric(x[,i])){
        for(j in 1:ncol(x)){
          if (!is.numeric(x[,j]) & names(x)[j]!=names(x)[i]){
            print(ggplot(data=x)+geom_bar(mapping=aes(x=x[,i],fill=x[,j]),position="dodge")+ labs(x=names(x)[i],fill=names(x)[j]))
          }
        }
      }
    }
}
F11(titanic)
```


Function 13
Write a function that do visualization of a dataset, i.e. the function should output a set of several graphs of the variables
```{r, include=TRUE}
F13 <- function(x){
  for (i in 1:ncol(x)){
      if (!is.numeric(x[,i])){
        for(j in 1:ncol(x)){
          if (!is.numeric(x[,j]) & names(x)[j]!=names(x)[i]){
            print(ggplot(data=x)+geom_bar(mapping=aes(x=x[,i],fill=x[,j]),position="dodge")+ labs(x=names(x)[i],fill=names(x)[j]))
          }
        }
      }else{
        for(j in 1:ncol(x)){
          if (!is.numeric(x[,j])){
            print(ggplot(data=x)+geom_density(mapping=aes(x=x[,i],color=x[,j]))+labs(x=names(x)[i],color=names(x)[j]))
          }
        }
      }
  }
}
F13(titanic)
```