---
title: "Assignment 10"
author: "Tavish Boyle"
date: "October 17, 2018"
output:
  word_document: default
  html_document: default
---

#1.	Download the adult dataset.  Change its extension to .csv


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#2.	Import to Rstudio setting the header = FALSE so that the first rows does not turn in to variable names.  This means:  df = read.csv("adult.csv", header = FALSE)
```{r, include=TRUE}
df = read.csv("C:/Users/student/Documents/Senior Year/Statistical Analysis with R/Data/adult.csv", header = FALSE, na.strings = c(" ?", "?", "."))
```

#3.	Change the variables' names to the appropriate ones.  You can use function names(df) to access the variables names. 

```{r, include=TRUE}
names(df)[1] <- "age"
names(df)[2] <- "workclass"
names(df)[3] <- "fnlwgt"
names(df)[4] <- "education"
names(df)[5] <- "education-num"
names(df)[6] <- "marital-status"
names(df)[7] <- "occupation"
names(df)[8] <- "relationship"
names(df)[9] <- "race"
names(df)[10] <- "sex"
names(df)[11] <- "capital-gain"
names(df)[12] <- "capital-loss"
names(df)[13] <- "hours-per-week"
names(df)[14] <- "native-country"
names(df)[15] <- "class"
```


#4.	Identify all the missing values.  Notice that missing values can be in the form of "NA", "?", "." and so on  
```{r, include=TRUE}
sum(is.na(df))
```

#5.	Replacing numeric missing values by the mean of the corresponding variables, and categorical missing values by the mode of the corresponding variables.
```{r, include=TRUE}
  for(i in 1:ncol(df))
     if (is.factor(df[,i]))
        levels=unique(df[,i])
        df[,i][is.na(df[,i])]=levels[which.max(tabulate(match(df[,i], levels)))]
      
sum(is.na(df))
```
#6.	Perform data visualization on the data.  You should include, at least:
#-Bar charts of two categorical variables
#-	Density curves of numeric variables on the categories of categorical variables.  For example: plotting density curves of Age on female and male. 
```{r, include = TRUE}
library(ggplot2)
ggplot(df, aes(x=relationship)) + geom_bar(aes(fill = sex), position = 'fill')

ggplot(df, aes(x = sex)) + geom_density(aes(alpha = race))
```

#7.	Partition the data into training and testing data with ratio train : test = 60 : 40

```{r, include=TRUE}
names(df)[15] = "target"
library(caret)
library(lattice)
splitIndex <- createDataPartition(df$target, p = .60, list = FALSE, times = 1)
train <- df[splitIndex,]
test <- df[-splitIndex,]
```

#8.	Build a decision tree model on the training data and test the model on the testing data
```{r, include=TRUE}
library(rpart)
 mytree.df <- rpart(target ~ ., data = train, method = "class")
```

#9.	Report the misclassification rate and confusion table of the tree model
```{r, include=TRUE}
pred.df <- predict(mytree.df,test, type = "class")
cm.df <- confusionMatrix(data = pred.df, reference = test$target)

cm.df
```

#10.	Print out the tree
```{r, include=TRUE}
library(rattle)

fancyRpartPlot(mytree.df)
```


#11.	Build tree model using caret package and report the confusion table 
#Could not get this to run properly. 
```{r}
library(caret)
model.df <- train(target~.,data = train, method = "rpart")
pred2.df <- predict(model.df, test)

cm2 <- confusionMatrix(pred2.df, test$target)

cm2
```

#12.	Write a function, taking the input of training data percentage, to automate 7-10.  Apply the function with training data = 70%; 80%; 90%
```{r, include=TRUE}
model.function <- function (train.amount) {
  
  names(df)[15] = "target"
library(caret)
  splitIndex <- createDataPartition(df$target, p = train.amount , list = FALSE, times = 1)
  train <- df[splitIndex,]
  test <- df[-splitIndex,]
  
  library(rpart)
  mytree.df <- rpart(target ~ ., data = train, method = "class")

  pred.df <- predict(mytree.df,test, type = "class")
  cm.df <- confusionMatrix(data = pred.df, reference = test$target)
  
  library(rattle)
  fancyRpartPlot(mytree.df)
  
}
model.function(.7)
model.function(.8)
model.function(.9)
```


