---
title: "Assignment 11"
author: "Tavish Boyle"
date: "October 22, 2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(lattice)
library(caret)
library(ranger)
library(e1071)

adult=read.csv("C:/Users/student/Documents/Senior Year/Statistical Analysis with R/Data/adult.csv", header = FALSE)

colnames(adult) <- c("age", "workclass", "fnlwgt", "education", "education_num", "marital_status", "occupation", "relationship", "race", "sex", "capital_gain", "capital_loss", "hours_per_week", "native_country", "salary")

```

##1.	Replace all the missing values by NA 
```{r}
adult[adult==' ?']=NA
sum(is.na(adult))
```

##2.	Replace the categorical missing values by the mode of the corresponding variables
```{r}
missingval=function(x) {
  for(i in 1:ncol(x)){
    if(!is.numeric(x[,i])){
      levels=unique(x[,i])
      x[,i][is.na(x[,i])]=levels[which.max(tabulate(match(x, levels)))]
    }
  }
  return(x)
}
adult2=missingval(adult)
sum(is.na(adult2))
```

##3.	Replace the numeric missing values by the median using caret

```{r}
num_miss=preProcess(adult2, method='medianImpute')
adult6=predict(num_miss, newdata = adult2)
sum(is.na(adult6))
```



##4.	The variable "native country" has 16 categories. Group the categories of this variable into 5 categories.  Do the same for all categorical variables with more than 10 categories. 
```{r}
adult6$native_country=as.character(adult6$native_country)
adult6$education=as.character(adult6$education)
adult6$occupation=as.character(adult6$occupation)

adult6$native_country[adult6$native_country==" United-States"|adult6$native_country==" Canada"|adult6$native_country==" Mexico"|adult6$native_country==" Cuba"|adult6$native_country==" Dominican-Republic"|adult6$native_country==" El-Salvador"|adult6$native_country==" Guatemala"|adult6$native_country==" Haiti"|adult6$native_country==" Honduras"|adult6$native_country==" Jamaica"|adult6$native_country==" Nicaragua"]="N. America"

adult6$native_country[adult6$native_country==" Cambodia"|adult6$native_country==" China"|adult6$native_country==" Hong"|adult6$native_country==" India"|adult6$native_country==" Iran"|adult6$native_country==" Japan"|adult6$native_country==" Laos"|adult6$native_country==" Philippines"|adult6$native_country==" Taiwan"|adult6$native_country==" Thailand"|adult6$native_country==" Vietnam"]="Asia"

adult6$native_country[adult6$native_country==" England"|adult6$native_country==" France"|adult6$native_country==" Germany"|adult6$native_country==" Greece"|adult6$native_country==" Holand-Netherlands"|adult6$native_country==" Hungary"|adult6$native_country==" Ireland"|adult6$native_country==" Italy"|adult6$native_country==" Poland"|adult6$native_country==" Portugal"|adult6$native_country==" Scotland"|adult6$native_country==" Yugoslavia"]="Euro"

adult6$native_country[adult6$native_country==" South"|adult6$native_country==" Outlying-US(Guam-USVI-etc)"|adult6$native_country==" Puerto-Rico"]="Other"

adult6$native_country[adult6$native_country==" Columbia"|adult6$native_country==" Ecuador"|adult6$native_country==" Peru"|adult6$native_country==" Trinadad&Tobago"]="S. America"

adult6$education[adult6$education==" HS-grad"|adult6$education==" 12th"|adult6$education==" 11th"|adult6$education==" 10th"|adult6$education==" 9th"]="H. School"

adult6$education[adult6$education==" Some-college"|adult6$education==" Assoc-acdm"|adult6$education==" Assoc-voc"]="Some College"

adult6$education[adult6$education==" 5th-6th"|adult6$education==" 7th-8th"]="Mid. School"

adult6$education[adult6$education==" Preschool"|adult6$education==" 1st-4th"]="Ele. School"

adult6$education[adult6$education==" Bachelors"|adult6$education==" Masters"|adult6$education==" Doctorate"|adult6$education==" Prof-school"]="Degree"


adult6$occupation[adult6$occupation==" Armed-Forces"|adult6$occupation==" Protective-serv"]="Defence"
adult6$occupation[adult6$occupation==" Craft-repair"|adult6$occupation==" Farming-fishing"|adult6$occupation==" Handlers-cleaners"|adult6$occupation==" Machine-op-inspct"|adult6$occupation==" Transport-moving"]="Physical"
adult6$occupation[adult6$occupation==" Adm-clerical"|adult6$occupation==" Exec-managerial"|adult6$occupation==" Sales"|adult6$occupation==" Tech-support"]="Office"
adult6$occupation[adult6$occupation==" Prof-specialty"]="Teaching"
adult6$occupation[adult6$occupation==" Other-service"|adult6$occupation==" Priv-house-serv"]="Misc"

adult6$native_country=as.factor(adult6$native_country)
adult6$education=as.factor(adult6$education)
adult6$occupation=as.factor(adult6$occupation)

levels(adult6$native_country)
levels(adult6$education)
levels(adult6$occupation)

```

##5.Encoding categorical variable using one hot encoding (dummy encoding) 
```{r}
dummies_model <- dummyVars(salary ~ ., data=adult)
adultData_mat <- predict(dummies_model, newdata = adult)

adultData <- data.frame(adultData_mat)
adultData$salary <- adult$salary

```

##6.	Scale and center the data
```{r}
preProcess_missingdata_model <- preProcess(adult6, method= c("center", "scale"))
preProcess_missingdata_model

```

##7.	Split the data into training (70%) and testing (30%) with the seeding set to be 2018 [including the code: set.seed(2018)]. Build a decision tree and report the accuracy and balanced accuracy
```{r}
splitIndex <- createDataPartition(adult6$salary, p = .70, list = FALSE, times = 1)
train <- adult6[ splitIndex,]
test <- adult6[-splitIndex,]
```

##8.	Build a random forest using the ranger package.  Report the accuracy and balanced accuracy
```{r}
library(ranger)
model = ranger(salary ~., data = train)
pred3  = predict(model, data = test)$predictions
cm3 = confusionMatrix(pred3, test$salary, positive=" >50K")
print(cm3)

```

##9.	Redo 3 with missing values being replaced by the mean.  Rebuild the model and report the models' performances (the accuracy and balanced accuracy). 
```{r}
#3
miss_num=function(x){
  for (i in 1:ncol(x)){
    if (is.numeric(x[,i])){
      x[,i][is.na(x[,i])]=mean(x[,i], na.rm=TRUE)
    }
  }
  return(x)
}
adult4=miss_num(adult2)
sum(is.na(adult4))



#4
adult4$native_country=as.character(adult4$native_country)
adult4$education=as.character(adult4$education)
adult4$occupation=as.character(adult4$occupation)

adult4$native_country[adult4$native_country==" United-States"|adult4$native_country==" Canada"|adult4$native_country==" Mexico"|adult4$native_country==" Cuba"|adult4$native_country==" Dominican-Republic"|adult4$native_country==" El-Salvador"|adult4$native_country==" Guatemala"|adult4$native_country==" Haiti"|adult4$native_country==" Honduras"|adult4$native_country==" Jamaica"|adult4$native_country==" Nicaragua"]="N. America"

adult4$native_country[adult4$native_country==" Cambodia"|adult4$native_country==" China"|adult4$native_country==" Hong"|adult4$native_country==" India"|adult4$native_country==" Iran"|adult4$native_country==" Japan"|adult4$native_country==" Laos"|adult4$native_country==" Philippines"|adult4$native_country==" Taiwan"|adult4$native_country==" Thailand"|adult4$native_country==" Vietnam"]="Asia"

adult4$native_country[adult4$native_country==" England"|adult4$native_country==" France"|adult4$native_country==" Germany"|adult4$native_country==" Greece"|adult4$native_country==" Holand-Netherlands"|adult4$native_country==" Hungary"|adult4$native_country==" Ireland"|adult4$native_country==" Italy"|adult4$native_country==" Poland"|adult4$native_country==" Portugal"|adult4$native_country==" Scotland"|adult4$native_country==" Yugoslavia"]="Euro"

adult4$native_country[adult4$native_country==" South"|adult4$native_country==" Outlying-US(Guam-USVI-etc)"|adult4$native_country==" Puerto-Rico"]="Other"

adult4$native_country[adult4$native_country==" Columbia"|adult4$native_country==" Ecuador"|adult4$native_country==" Peru"|adult4$native_country==" Trinadad&Tobago"]="S. America"

adult4$education[adult4$education==" HS-grad"|adult4$education==" 12th"|adult4$education==" 11th"|adult4$education==" 10th"|adult4$education==" 9th"]="H. School"

adult4$education[adult4$education==" Some-college"|adult4$education==" Assoc-acdm"|adult4$education==" Assoc-voc"]="Some College"

adult4$education[adult4$education==" 5th-6th"|adult4$education==" 7th-8th"]="Mid. School"

adult4$education[adult4$education==" Preschool"|adult4$education==" 1st-4th"]="Ele. School"

adult4$education[adult4$education==" Bachelors"|adult4$education==" Masters"|adult4$education==" Doctorate"|adult4$education==" Prof-school"]="Degree"


adult4$occupation[adult4$occupation==" Armed-Forces"|adult4$occupation==" Protective-serv"]="Defence"
adult4$occupation[adult4$occupation==" Craft-repair"|adult4$occupation==" Farming-fishing"|adult4$occupation==" Handlers-cleaners"|adult4$occupation==" Machine-op-inspct"|adult4$occupation==" Transport-moving"]="Physical"
adult4$occupation[adult4$occupation==" Adm-clerical"|adult4$occupation==" Exec-managerial"|adult4$occupation==" Sales"|adult4$occupation==" Tech-support"]="Office"
adult4$occupation[adult4$occupation==" Prof-specialty"]="Teaching"
adult4$occupation[adult4$occupation==" Other-service"|adult4$occupation==" Priv-house-serv"]="Misc"

adult4$native_country=as.factor(adult4$native_country)
adult4$education=as.factor(adult4$education)
adult4$occupation=as.factor(adult4$occupation)

levels(adult4$native_country)
levels(adult4$education)
levels(adult4$occupation)


#5
dummies_model <- dummyVars(salary ~ ., data=adult)
adultData_mat <- predict(dummies_model, newdata = adult)

adultData <- data.frame(adultData_mat)
adultData$salary <- adult$salary



#6
preProcess_missingdata_model <- preProcess(adult4, method= c("center", "scale"))
preProcess_missingdata_model


#7
splitIndex <- createDataPartition(adult4$salary, p = .70, list = FALSE, times = 1)
train <- adult4[ splitIndex,]
test <- adult4[-splitIndex,]


#8
library(ranger)
model = ranger(salary ~., data = train)
pred4  = predict(model, data = test)$predictions
cm4 = confusionMatrix(pred4, test$salary, positive=" >50K")
print(cm4)

```

##10.	Redo 3 with missing values being replaced by the "knn" imputation method.  Rebuild the model and report the models' performances (the accuracy and balanced accuracy)
```{r}
#3
num_miss=preProcess(adult2, method='knnImpute')
adult5=predict(num_miss, newdata = adult2)
sum(is.na(adult5))


#4
adult5$native_country=as.character(adult5$native_country)
adult5$education=as.character(adult5$education)
adult5$occupation=as.character(adult5$occupation)

adult5$native_country[adult5$native_country==" United-States"|adult5$native_country==" Canada"|adult5$native_country==" Mexico"|adult5$native_country==" Cuba"|adult5$native_country==" Dominican-Republic"|adult5$native_country==" El-Salvador"|adult5$native_country==" Guatemala"|adult5$native_country==" Haiti"|adult5$native_country==" Honduras"|adult5$native_country==" Jamaica"|adult5$native_country==" Nicaragua"]="N. America"

adult5$native_country[adult5$native_country==" Cambodia"|adult5$native_country==" China"|adult5$native_country==" Hong"|adult5$native_country==" India"|adult5$native_country==" Iran"|adult5$native_country==" Japan"|adult5$native_country==" Laos"|adult5$native_country==" Philippines"|adult5$native_country==" Taiwan"|adult5$native_country==" Thailand"|adult5$native_country==" Vietnam"]="Asia"

adult5$native_country[adult5$native_country==" England"|adult5$native_country==" France"|adult5$native_country==" Germany"|adult5$native_country==" Greece"|adult5$native_country==" Holand-Netherlands"|adult5$native_country==" Hungary"|adult5$native_country==" Ireland"|adult5$native_country==" Italy"|adult5$native_country==" Poland"|adult5$native_country==" Portugal"|adult5$native_country==" Scotland"|adult5$native_country==" Yugoslavia"]="Euro"

adult5$native_country[adult5$native_country==" South"|adult5$native_country==" Outlying-US(Guam-USVI-etc)"|adult5$native_country==" Puerto-Rico"]="Other"

adult5$native_country[adult5$native_country==" Columbia"|adult5$native_country==" Ecuador"|adult5$native_country==" Peru"|adult5$native_country==" Trinadad&Tobago"]="S. America"

adult5$education[adult5$education==" HS-grad"|adult5$education==" 12th"|adult5$education==" 11th"|adult5$education==" 10th"|adult5$education==" 9th"]="H. School"

adult5$education[adult5$education==" Some-college"|adult5$education==" Assoc-acdm"|adult5$education==" Assoc-voc"]="Some College"

adult5$education[adult5$education==" 5th-6th"|adult5$education==" 7th-8th"]="Mid. School"

adult5$education[adult5$education==" Preschool"|adult5$education==" 1st-4th"]="Ele. School"

adult5$education[adult5$education==" Bachelors"|adult5$education==" Masters"|adult5$education==" Doctorate"|adult5$education==" Prof-school"]="Degree"


adult5$occupation[adult5$occupation==" Armed-Forces"|adult5$occupation==" Protective-serv"]="Defence"
adult5$occupation[adult5$occupation==" Craft-repair"|adult5$occupation==" Farming-fishing"|adult5$occupation==" Handlers-cleaners"|adult5$occupation==" Machine-op-inspct"|adult5$occupation==" Transport-moving"]="Physical"
adult5$occupation[adult5$occupation==" Adm-clerical"|adult5$occupation==" Exec-managerial"|adult5$occupation==" Sales"|adult5$occupation==" Tech-support"]="Office"
adult5$occupation[adult5$occupation==" Prof-specialty"]="Teaching"
adult5$occupation[adult5$occupation==" Other-service"|adult5$occupation==" Priv-house-serv"]="Misc"

adult5$native_country=as.factor(adult5$native_country)
adult5$education=as.factor(adult5$education)
adult5$occupation=as.factor(adult5$occupation)

levels(adult5$native_country)
levels(adult5$education)
levels(adult5$occupation)


#5
dummies_model <- dummyVars(salary ~ ., data=adult5)
adult5Data_mat <- predict(dummies_model, newdata = adult5)

adult5Data <- data.frame(adult5Data_mat)
adult5Data$salary <- adult5$salary


#6
preProcess_missingdata_model <- preProcess(adult5, method= c("center", "scale"))
preProcess_missingdata_model

#7
splitIndex <- createDataPartition(adult5$salary, p = .70, list = FALSE, times = 1)
train <- adult5[ splitIndex,]
test <- adult5[-splitIndex,]


#8
library(ranger)
model = ranger(salary ~., data = train)
pred5  = predict(model, data = test)$predictions
cm5 = confusionMatrix(pred5, test$salary, positive=" >50K")
print(cm5)

```

##11.	Redo 6 with only scaling and centering the non-encoded variables. Rebuild the model and report the models' performances (the accuracy and balanced accuracy)
```{r}
num_miss=preProcess(adult2, method='medianImpute')
adult6=predict(num_miss, newdata = adult2)
sum(is.na(adult6))


adult6$native_country=as.character(adult6$native_country)
adult6$education=as.character(adult6$education)
adult6$occupation=as.character(adult6$occupation)

adult6$native_country[adult6$native_country==" United-States"|adult6$native_country==" Canada"|adult6$native_country==" Mexico"|adult6$native_country==" Cuba"|adult6$native_country==" Dominican-Republic"|adult6$native_country==" El-Salvador"|adult6$native_country==" Guatemala"|adult6$native_country==" Haiti"|adult6$native_country==" Honduras"|adult6$native_country==" Jamaica"|adult6$native_country==" Nicaragua"]="N. America"

adult6$native_country[adult6$native_country==" Cambodia"|adult6$native_country==" China"|adult6$native_country==" Hong"|adult6$native_country==" India"|adult6$native_country==" Iran"|adult6$native_country==" Japan"|adult6$native_country==" Laos"|adult6$native_country==" Philippines"|adult6$native_country==" Taiwan"|adult6$native_country==" Thailand"|adult6$native_country==" Vietnam"]="Asia"

adult6$native_country[adult6$native_country==" England"|adult6$native_country==" France"|adult6$native_country==" Germany"|adult6$native_country==" Greece"|adult6$native_country==" Holand-Netherlands"|adult6$native_country==" Hungary"|adult6$native_country==" Ireland"|adult6$native_country==" Italy"|adult6$native_country==" Poland"|adult6$native_country==" Portugal"|adult6$native_country==" Scotland"|adult6$native_country==" Yugoslavia"]="Euro"

adult6$native_country[adult6$native_country==" South"|adult6$native_country==" Outlying-US(Guam-USVI-etc)"|adult6$native_country==" Puerto-Rico"]="Other"

adult6$native_country[adult6$native_country==" Columbia"|adult6$native_country==" Ecuador"|adult6$native_country==" Peru"|adult6$native_country==" Trinadad&Tobago"]="S. America"

adult6$education[adult6$education==" HS-grad"|adult6$education==" 12th"|adult6$education==" 11th"|adult6$education==" 10th"|adult6$education==" 9th"]="H. School"

adult6$education[adult6$education==" Some-college"|adult6$education==" Assoc-acdm"|adult6$education==" Assoc-voc"]="Some College"

adult6$education[adult6$education==" 5th-6th"|adult6$education==" 7th-8th"]="Mid. School"

adult6$education[adult6$education==" Preschool"|adult6$education==" 1st-4th"]="Ele. School"

adult6$education[adult6$education==" Bachelors"|adult6$education==" Masters"|adult6$education==" Doctorate"|adult6$education==" Prof-school"]="Degree"


adult6$occupation[adult6$occupation==" Armed-Forces"|adult6$occupation==" Protective-serv"]="Defence"
adult6$occupation[adult6$occupation==" Craft-repair"|adult6$occupation==" Farming-fishing"|adult6$occupation==" Handlers-cleaners"|adult6$occupation==" Machine-op-inspct"|adult6$occupation==" Transport-moving"]="Physical"
adult6$occupation[adult6$occupation==" Adm-clerical"|adult6$occupation==" Exec-managerial"|adult6$occupation==" Sales"|adult6$occupation==" Tech-support"]="Office"
adult6$occupation[adult6$occupation==" Prof-specialty"]="Teaching"
adult6$occupation[adult6$occupation==" Other-service"|adult6$occupation==" Priv-house-serv"]="Misc"

adult6$native_country=as.factor(adult6$native_country)
adult6$education=as.factor(adult6$education)
adult6$occupation=as.factor(adult6$occupation)

levels(adult6$native_country)
levels(adult6$education)
levels(adult6$occupation)

dummies_model <- dummyVars(salary ~ ., data = adult6)
trainData_mat <- predict(dummies_model, newdata = adult6)

trainData <- data.frame(trainData_mat)
trainData$salary <- adult6$salary


numdf <- c()
catdf <- c()
  for(i in 1:ncol(trainData)) {
    if(is.numeric(trainData[,i])) {
      numdf <- c(numdf, names(trainData)[i])
    }
    else {
      catdf <- c(catdf, names(trainData)[i])
    }
  }
  numdata <- trainData[numdf]
  catdata <- trainData[catdf]


trainData2 <- preProcess(numdata, method= c("center", "scale"))
trainData2


#7
names(adult6)[15] = "salary"
set.seed(2018)
splitIndex <- createDataPartition(adult6$salary, p = .70, list = FALSE, times = 1)
train <- adult6[ splitIndex,]
test <- adult6[-splitIndex,]



#8
library(ranger)
model = ranger(salary ~., data = train)
pred10  = predict(model, data = test)$predictions
cm10 = confusionMatrix(pred10, test$salary, positive=" >50K")
print(cm10)

```

##12.	Redo 5 with a different encoding method. Rebuild the model and report the models' performances (the accuracy and balanced accuracy). 
```{r}

adult$workclass = as.numeric(adult$workclass)
adult$education = as.numeric(adult$education)
adult$marital_status = as.numeric(adult$marital_status)
adult$occupation = as.numeric(adult$occupation)
adult$relationship = as.numeric(adult$relationship)
adult$race = as.numeric(adult$race)
adult$sex = as.numeric(adult$sex)
adult$native_country = as.numeric(adult$native_country)

#1
adult[adult==' ?']=NA
sum(is.na(adult))

#2
missingval=function(x) {
  for(i in 1:ncol(x)){
    if(!is.numeric(x[,i])){
      levels=unique(x[,i])
      x[,i][is.na(x[,i])]=levels[which.max(tabulate(match(x, levels)))]
    }
  }
  return(x)
}
adult2=missingval(adult)
sum(is.na(adult2))

#3
num_miss=preProcess(adult2, method='medianImpute')
adult7=predict(num_miss, newdata = adult2)
sum(is.na(adult7))

#5

dummies_model <- dummyVars(salary ~ ., data=adult)
adultData_mat <- predict(dummies_model, newdata = adult)

adultData <- data.frame(adultData_mat)
adultData$salary <- adult$salary

#6
preProcess_missingdata_model <- preProcess(adult7, method= c("center", "scale"))
preProcess_missingdata_model

#7
splitIndex <- createDataPartition(adult7$salary, p = .70, list = FALSE, times = 1)
train <- adult7[ splitIndex,]
test <- adult7[-splitIndex,]


#8
library(ranger)
model = ranger(salary ~., data = train)
pred8  = predict(model, data = test)$predictions
cm8 = confusionMatrix(pred8, test$salary, positive=" >50K")
print(cm8)
```


##13.	Skip 4 to rebuild the models.  Do you find any improvement in the models' performances?  
```{r}
adult[adult==' ?']=NA
sum(is.na(adult))

#2
missingval=function(x) {
  for(i in 1:ncol(x)){
    if(!is.numeric(x[,i])){
      levels=unique(x[,i])
      x[,i][is.na(x[,i])]=levels[which.max(tabulate(match(x, levels)))]
    }
  }
  return(x)
}
adult2=missingval(adult)
sum(is.na(adult2))

#3
num_miss=preProcess(adult2, method='medianImpute')
adult8=predict(num_miss, newdata = adult2)
sum(is.na(adult8))

#5
dummies_model <- dummyVars(salary ~ ., data=adult8)
adult8Data_mat <- predict(dummies_model, newdata = adult8)

adult8Data <- data.frame(adult8Data_mat)
adult8Data$salary <- adult8$salary


#6
preProcess_missingdata_model <- preProcess(adult8, method= c("center", "scale"))
preProcess_missingdata_model

#7
splitIndex <- createDataPartition(adult8$salary, p = .70, list = FALSE, times = 1)
train <- adult8[ splitIndex,]
test <- adult8[-splitIndex,]


#8
library(ranger)
model = ranger(salary ~., data = train)
pred6  = predict(model, data = test)$predictions
cm6 = confusionMatrix(pred6, test$salary, positive=" >50K")
print(cm6)
```
