---
title: "Assignment 18"
author: "Tavish Boyle"
date: "December 5, 2018"
output: word_document
---

## Questions 1-6
```{r setup, include=TRUE}

df = read.csv("C:/Users/student/Documents/Senior Year/Statistical Analysis with R/Data/hmeq.csv")
summary(df)
d <- df[complete.cases(df),]
nrow(d)
d$BAD = as.factor(d$BAD)
names(d)[1]
names(d)[1] = 'target'
table(d$target)

library(caret)
library(ggplot2)
library(lattice)
library(ranger)
splitIndex <-createDataPartition(d$target, p=.70, list = FALSE, times = 1)
train <-d[splitIndex,]
test <- d[-splitIndex,]





model = ranger(target~., data = train)
pred = predict(model, data = test)$predictions
cm=confusionMatrix(pred, test$target, positive = "1")
cm

table(d$target)

prop.table(table(d$target))

train1 = train[train$target == "1",]
nrow(train1)



table(train1$target)

train0 = train[train$target == '0',]
table(train0$target)

sample(1:10, 3)

sample(nrow(train0), 217)

sample(nrow(train0), nrow(train1))


n=nrow(train0)
n1 = nrow(train1)
train0[sample(1:n, n1, replace = TRUE),]

train00=train0[sample(1:n,n1, replace = TRUE),]

train_under=rbind(train00,train1)
model_under=ranger(target~.,data=train_under)
pred_under=predict(model_under,data=test)$predictions
cm_under=confusionMatrix(pred_under,test$target,positive="1")
cm_under$overall['Accuracy']
cm_under$byClass['Balanced Accuracy']


table(train_under$target)
##write.csv(train_under, 'train_under.csv')

```

## Question 7
```{r}

summary(df)
d <- df[complete.cases(df),]
d$BAD = as.factor(d$BAD)
nrow(d)
names(d)[1]
names(d)[1] = 'target'
table(d$target)

splitIndex <-createDataPartition(d$target, p=.70, list = FALSE, times = 1)
train <-d[splitIndex,]
test <- d[-splitIndex,]





model = ranger(target~., data = train)
pred = predict(model, data = test)$predictions
cm=confusionMatrix(pred, test$target, positive = "1")
cm

table(d$target)

prop.table(table(d$target))

train1 = train[train$target == "1",]
nrow(train1)

table(train1$target)

train0 = train[train$target == '0',]
table(train0$target)

sample(1:10, 3)

sample(nrow(train0), 217)

sample(nrow(train0), nrow(train1))


n0=nrow(train0)
n1 = nrow(train1)
train11 = train1[sample(1:n1, n0, replace = TRUE),]


train_over=rbind(train11,train0)
model_over=ranger(target~.,data=train_over)
pred_over=predict(model_over,data=test)$predictions
cm_over=confusionMatrix(pred_over,test$target,positive="1")
cm_over$overall['Accuracy']
cm_over$byClass['Balanced Accuracy']


table(train_over$target)
##write.csv(train_under, 'train_over.csv')
```

## Question 8	Write a function called Quick_bal.  This function takes
##-	Inputs:  a dataset argument with a target variable named target; a method argument specifying undersampling or oversampling
##-	Output: a dataset with a balanced target. 

```{r}
Quick_bal <- function(x,m){
  names(x)[1] = 'target'
  splitIndex <- createDataPartition(x$target, p=.70, list = FALSE, times = 1)
  train <-d[splitIndex,]
  test <- d[-splitIndex,]

    if(m==1){
      train1 = train[train$target == "1",]
      train0 = train[train$target == "0",]
      n0 = nrow(train0)
      n1 = nrow(train1)
      train00 = train0[sample(1:n0, n1),]
      train_under = rbind(train00, train1)
    }
      if(m == 2){
        train1 = train[train$target == "1",]
        train0 = train[train$target == "0",]
        n0 = nrow(train0)
        n1 = nrow(train1)
        train11 = train1[sample(1:n1, n0, replace = TRUE),]
        train_over = rbind(train0, train11)
      }
}
```

## Question 9	Put 6 and 7 into a function called Quick_model_bal. This function takes
##-	Inputs: a dataset with a target variable named target 
##-	Output: the balanced accuracies of random forest with undersampling and oversampling being applied on the training dataset. 
```{r}
Quick_model_bal <- function(x,m){
  names(x)[1] = 'target'
  splitIndex <- createDataPartition(x$target, p=.70, list = FALSE, times = 1)
  train <-d[splitIndex,]
  test <- d[-splitIndex,]
  
  print("Undersample:")
  model_under=ranger(target~.,data=train_under)
pred_under=predict(model_under,data=test)$predictions
cm_under=confusionMatrix(pred_under,test$target,positive="1")
cm_under$overall['Accuracy']
cm_under$byClass['Balanced Accuracy']
  
  
  print("Oversample:")
  train_over=rbind(train11,train0)
model_over=ranger(target~.,data=train_over)
pred_over=predict(model_over,data=test)$predictions
cm_over=confusionMatrix(pred_over,test$target,positive="1")
cm_over$overall['Accuracy']
cm_over$byClass['Balanced Accuracy']
}
```


